Using demographic accounts to enhance official population statistics

### 2021-11-02 article.tex

A slightly more complicated account is
\begin{center}
\includegraphics[width=0.5\textwidth]{figures_accounts/fig_account_region.pdf}
\end{center}
Rather than treating the population as a completely homogenous, the account distinguishes two groups. If  whole,  
We include migration between the two regions in the account, but for this artificial example we assume that there are no births, deaths, or migrations to and from the rest of the world. We apply the demographic accounting identities separately to each region,
\begin{align*}
    11 - 15 & = -5 + 1 \quad \text{(A)}\\
    14 - 10 & = -1 + 5 \quad \text{(B)}.
\end{align*}
When the accounting identities for each region are satisfied, the accounting identity for the system is a whole is necessarily satisfied as well.

One particular type of disaggregation has a special status within demography: disaggregation by age. Age is special because it is such a strong predictor of demographic behavior. The mortality and migration rates of 20 year olds and 90 year olds, for instance, differ by orders of magnitude. Age is also special because movement between age groups is deterministic: everyone is born into the lowest age group, and then moves upwards at the rate of one age group per period.

Publishers of age-structured demographic data typically use some variant of the following scheme,
\begin{center}
\includegraphics[width=0.5\textwidth]{figures_accounts/fig_account_withage2.pdf}
\end{center}
Births = lowest. But difficult, cannot look across age groups. Demographic accounting only at population level.

>>>>

A slightly more complicated example is
\begin{center}
\includegraphics[width=0.5\textwidth]{figures_accounts/fig_account_region.pdf}
\end{center}
Rather than treating the population as a completely homogenous, this account recognizes two subpopulations, labelled ``A'' and ``B''. A and B might, for instance, denote geographical region, health status, or labour force status. In this example, for simplicity, we allow for movement between the two statuses, but no births, deaths, or other forms of entry or exit to the system as a whole. The demographic accounting identities can be applied separately to each group,
\begin{align*}
    11 - 15 & = -5 + 1 \quad \text{(A)}\\
    14 - 10 & = -1 + 5 \quad \text{(B)}.
\end{align*}
When the accounting identities for each group are satisfied, the accounting identity for the system is a whole is necessarily satisfied as well.

One particular type of disaggregation into subpopulations has a special status within demography: disaggregation by age group. Age is special because it is such a strong predictor of demographic behavior. The mortality and migration rates of 20-24 year olds and 90-94 year olds, for instance, differ by several orders of magnitude. Age is also special because movement between age groups is deterministic. Everyone is born into the lowest age group, and then moves upwards at the rate of one age group per period.

Publishers of age-structured demographic data typically use some variant of the following scheme,
\begin{center}
\includegraphics[width=0.5\textwidth]{figures_accounts/fig_account_withage2.pdf}
\end{center}
Population stocks stocks and flows are all disaggregated by the same age groups, which, in the case above, are ages 0, 1, and 2+. (The account in question might apply to non-human animals, or use a unit of measurement that is longer than a year.) The final age group includes everone over a certain age. Births are disaggregated by the age of the mother, rather than the age of the child being born, which is always 0.

The problem with the standard scheme for presenting age-classified data is that it does allow the demographic accounting identities to be applied within age groups. If, for instance, we try to

Births = lowest. But difficult, cannot look across age groups. Demographic accounting only at population level.
